### 2014-10-29 contract-language.tex
Minor edit/comment
edits

\item \textbf{Causal consistency}: Causally consistent operations
  require to see a causally consistent snapshot of the object state,
  including the actions performed on the same session.  The latter
  requirement entails that if two operations $o_1$ and $o_2$ from the
  same session are applied to two different replicas $r_1$ and $r_2$,
  the second operation cannot be discharged until the effect of $o_1$ is
  merged with $o_2$ in both $r_1$ and $r_2$. The \cf{getBalance}
  operation requires causal consistency, as it requires the operations
  from the same session to be visible, which cannot be guaranteed under
  eventual consistency. We assume that causality is only tracked through
  operations on the same object; two operations in the same session but
  on different objects are considered causally unrelated under this
  definition. Stores typically avoid tracking causality across objects
  to mitigate overheads when causality tracking is unnecessary. The
  corresponding store semantics is captured by the contract $\ccc$
  defined below:

@@

Observe that out contract language does not incorporate real (i.e.,
wall-clock) time. Hence, the contract language cannot describe store
semantics based on real time such as recency or bounded-staleness
guarantees offered by certain stores~\cite{Pileus}.

@@

Our goal is to map application-level consistency constraints of
operations to appropriate store-level consistency guarantees capable
of satisfying constraints.  The ability to express both of them as
contracts in our contract language lets us compare and determine if
contract ($\cv_{op}$) of an operation ($op$) is weak enough to be
satisfied under a store consistency level identified by the contract
$\cv_{st}$. Towards this end, we define a binary \emph{weaker than}
relation for our contract language as following:

>>>>

\item \textbf{Causal consistency}: Causally consistent operations
  are required to see a causally consistent snapshot of the object state,
  including the actions performed on the same session.  The latter
  requirement entails that if two operations $o_1$ and $o_2$ from the
  same session are applied to two different replicas $r_1$ and $r_2$,
  the second operation cannot be discharged until the effect of $o_1$ is
  merged with $o_2$ in both $r_1$ and $r_2$. The \cf{getBalance}
  operation requires causal consistency, as it requires the operations
  from the same session to be visible, which cannot be guaranteed under
  eventual consistency. We assume that causality is only tracked through
  operations on the same object; two operations in the same session but
  on different objects are considered causally unrelated under this
  definition. Stores typically avoid tracking causality across objects
  to mitigate overheads when causality tracking is unnecessary. The
  corresponding store semantics is captured by the contract $\ccc$
  defined below:

@@

% SJ: not sure this is necessary, or appropriate here.
%% \noindent Observe that out contract language does not incorporate real (i.e.,
%% wall-clock) time. Hence, it cannot describe store
%% semantics such as recency or bounded-staleness
%% guarantees offered by certain stores~\cite{Pileus}.

@@

Our goal is to map application-level consistency constraints on
operations to appropriate store-level consistency guarantees capable
of satisfying these constraints.  The ability to express both of them as
contracts in our contract language lets us compare and determine if
contract ($\cv_{op}$) of an operation ($op$) is weak enough to be
satisfied under a store consistency level identified by the contract
$\cv_{st}$. Towards this end, we define a binary \emph{weaker than}
relation for our contract language as following:

### 2014-10-29 core-lang.tex
edits

\section{Meta-theory}
\label{sec:core-opsem}

In this section, we present the meta-theoretic result that certifies
the soundness of our classification-based contract enforcement
strategy.
% We show that any data store can correctly enforce $\scc$,
% $\ccc$, and $\ecc$ guarantees can satisfy application-level
% consistency constraints of all \name contracts.

Fig.~\ref{sem:oper} summarizes the artifacts needed to state our
meta-theoretic result. We formalize an abstract model of a replicated
data store as a tuple $\E;\Sigma$, where $\E$ denotes the execution
state, and $\Sigma$, the session soup, is the set of concurrent client
sessions interacting with the store. The execution state ($\E$) of a
data store is modeled as a tuple $\Exec$ where $\EffSoup$, called the
\emph{effect soup}, is a set of effects generated during the
execution, and $\visZ,\soZ \subseteq \EffSoup \times \EffSoup$ are
visibility and session order relations, respectively, witnessed over
generated effects. A session is modeled simply as a sequence of
replicated data type operations ($op$), tagged with the consistency
class ($\tau$) of their contracts (as determined by the contract
classification scheme). Our contract language and classification
scheme is independent of store semantics. Accordingly, we abstain from
making any specific assumptions about semantics of the store. However,
we do assume the existence of a reduction relation of form:

@@

\noindent The relation captures store's progress in execution (from
$\E$ to $\E'$) due to the application of an operation ($op$) from one
of the sessions in $\Sigma$ to a replica, generating a new effect
($\eff$).

@@

We now state the soundness of contract enforcement as following:

@@

The theorem states that if a data store correctly enforces $\scc$,
$\ccc$, and $\ecc$ guarantees in all executions, then contract
classification scheme can be used to satisfy any \name contract on
that data store. The proof of the theorem has been included in our
tech report~\cite{techrep}. Tech report also contains operational
semantics of our implementation of the data store, which concretizes
the reduction relation ($\xrightarrow{}$), along with the proof
that it correctly enforces all \name contracts.

>>>>

\section{Soundness}
\label{sec:core-opsem}

In this section, we present a soundness result that certifies
the correctness of our classification-based contract enforcement
strategy.
% We show that any data store can correctly enforce $\scc$,
% $\ccc$, and $\ecc$ guarantees can satisfy application-level
% consistency constraints of all \name contracts.

Fig.~\ref{sem:oper} summarizes the artifacts needed. We formalize an
abstract model of a replicated data store as a tuple $\E;\Sigma$, where $\E$
denotes the execution state, and $\Sigma$, the session soup, is the set of
concurrent client sessions interacting with the store. The execution state
($\E$) of a data store is modeled as a tuple $\Exec$ where $\EffSoup$,
called the \emph{effect soup}, is a set of effects generated during the
execution, and $\visZ,\soZ \subseteq \EffSoup \times \EffSoup$ are
visibility and session order relations, respectively, witnessed over
generated effects. A session is modeled simply as a sequence of replicated
data type operations ($op$), tagged with the consistency class ($\tau$) of
their contracts (as determined by the contract classification scheme). Our
contract language and classification scheme is independent of store
semantics. Accordingly, we avoid making any specific assumptions about the
semantics of the store. However, we do assume the existence of a reduction
relation of the form:

@@

\noindent The relation captures the store's execution progress (from
$\E$ to $\E'$) due to the application of an operation ($op$) from one
of the sessions in $\Sigma$ to a replica, generating a new effect
($\eff$).

@@

@@

The theorem states that if a data store correctly enforces $\scc$,
$\ccc$, and $\ecc$ guarantees in all executions, then contract
classification scheme can be used to satisfy any \name contract on
that data store. The proof of the theorem is available in the supplementary
material, which also contains an operational
semantics of our implementation of the data store that concretizes
the reduction relation ($\xrightarrow{}$), along with the proof
that it correctly enforces all \name contracts.
